---
title: "ANOVA_R_sameas_SPSS"
author: "liuc"
date: '2022-03-18'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ANOVA_R_sameas_SPSS

在做方差分析时，R里默认的contrasts和SPSS里面不一致, 现将其一致的结果整理如下：

http://www.statscanbefun.com/rblog/2015/8/27/ensuring-r-generates-the-same-anova-f-values-as-spss

https://medium.com/humansystemsdata/analysis-of-variance-showdown-r-vs-spss-f4e50234a94

he nature of the differences between SPSS and R becomes evident when there are an unequal number of participants across factorial ANOVA cells. There are a few simple steps that can be followed to ensure that R ANOVA values do indeed match those generated by SPSS. These steps involves using Type-III sums of squares for the ANOVA but there is more to it than that.


There are three things you need to do to ensure ANOVA F-values in R match those in SPSS. I will briefly list these three steps and then provide a more details description of each.

1. Set each independent variable as a factor
2. Set the default contrast to helmert
3. Conduct analysis using Type III Sums of Squares



```{r}
# We need to change the default contrast for unordered factors from "cont.treatment" to "contr.helmert". We do this with the command below

options(contrasts = c("contr.helmert", "contr.poly"))

df <- haven::read_sav('~/Downloads/两因素方差分析-SPSS教程-医咖会/16 双因素方差分析.sav')

dd <- aov(Index ~ gender * education, data = df)
car::Anova(dd, type = 'III')
           
```

```{r}
# same as dd, 用线性模型得到的结果
crf.lm <- lm(formula = Index ~ gender * education, data = df)

car::Anova(crf.lm)


```


## 重复测量方差分析
#### 单因素重复测量方差分析一般指单一的Time因素
主要讨论
#### 双因素重复测量方差分析

重复测量方差分析适用于结局变量为连续变量，且因变量符合正态分布 的数据。

比如常见的医学数据，对两组患者的在不同时间采样的某项定量数据，是最为常见的试验设计方式。一般因变量的测量为不同时期测量数据减去baseline数据作为结局变量。
对于AB两组各10个人的设计，time为within，group为between；对于10个人都进行两组不同的处理，则time和group都是within因素。

Time作为within-Subjects variables
between-Subject variables是指何种变量呢？group.
covariates 如何解释呢？

*单独效应：*
*主效应：*
*交互效应：*


```{r}
# 注意数据的category变量要变成factor
df <- haven::read_sav('./datasets/两因素重复测量的方差分析.sav')
df <- df %>% pivot_longer(cols = starts_with('t'), names_to = 'time', values_to = 'value') %>% 
  rstatix::convert_as_factor(group, id, time)

# 重复测量方差分析
aov_res <- aov(value ~ group * time + Error(id/time), data = df)

aov_res
summary(aov_res)

parameters::model_parameters(aov_res)
effectsize::eta_squared(aov_res)

# emmeans
grafify::posthoc_Levelwise(Model = aov_res,
                           Fixed_Factor = c("group"),
                           infer = c(TRUE, TRUE)
                           )


## use 
aov_res2 <- rstatix::anova_test(formula = value ~ group * time + Error(id/time),
                    data = df
                    )


```

*Repeated-measures ANOVA with the car package*
https://mspeekenbrink.github.io/sdam-r-companion/repeated-measures-anova.html

```{r}
library(car)
library(afex)

# car::Anova()需要宽数据，以下用afex包进行分析


# afex provides an abridged ANOVA table, where the Greenhouse-Geisser correction is automatically applied
afmod <- afex::aov_car(value ~ group * time + Error(id/time), 
                       data = df)
afmod
summary(afmod)
afex::nice(afmod, es="pes", correction = "none")

em_version <- emmeans::emmeans(afmod, specs = ~ group)
em_version
emmeans::contrast(em_version, method=list("group1 - group2" = c(1,-1)))

grafify::posthoc_Levelwise(Model = afmod,
                           Fixed_Factor = c("group"),
                           infer = c(TRUE, TRUE)
                           )

```







