---
title: "CANONICAL CORRELATION ANALYSIS"
author: "liuc"
date: "1/25/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CANONICAL CORRELATION ANALYSIS

典型相关分析 <https://stats.oarc.ucla.edu/r/dae/canonical-correlation-analysis/>

Canonical correlation analysis is used to identify and measure the associations among two sets of variables. Canonical correlation is appropriate in the same situations where multiple regression would be, but where are there are multiple intercorrelated outcome variables. Canonical correlation analysis determines a set of canonical variates, orthogonal linear combinations of the variables within each set that best explain the variability both within and between sets.

典型相关可以理解为一组变量和另外一组变量间的关系的方法。比如我们可以把多重线性回归看成是一个变量和一组变量间关系的描述，而CCA则是一组和一组，即是有多个因变量。这些一组变量间一般会归属于一类，nest。 根据变量间的关系，寻找少数几个关系简单的综合变量对，替代关系复杂的实际观察变量。这些综合变量被称为‘典型变量’。

*Examples of canonical correlation analysis* Example 1. A researcher has collected data on three psychological variables, four academic variables (standardized test scores) and gender for 600 college freshman. She is interested in how the set of psychological variables relates to the academic variables and gender. In particular, the researcher is interested in how many dimensions (canonical variables) are necessary to understand the association between the two sets of variables.

Example 2. A researcher is interested in exploring associations among factors from two multidimensional personality tests, the MMPI and the NEO. She is interested in what dimensions are common between the tests and how much variance is shared between them. She is specifically interested in finding whether the neuroticism dimension from the NEO can account for a substantial amount of shared variance between the two tests.

```{r}
require(ggplot2)
require(GGally)
require(CCA)
require(CCP)
```

```{r}
# 首先还是对数据的探索和条件的满足，不过此处只考虑数据输入格式和可以解决的问题。
mm <- read.csv("./datasets/mmreg.csv")
colnames(mm) <- c("Control", "Concept", "Motivation", "Read", "Write", "Math", 
    "Science", "Sex")
# 典型相关分析的模型假定
# 最好是定量数据、亦可为等级资料；满足多元正态分布；两组变量间为线性关系


# 两组相关变量间的关系，对于多组变量间的相关性？
psych <- mm[, 1:3]
acad <- mm[, 4:8]

# cancor()
cc1 <- CCA::cc(psych, acad)
cc1[3:4]


# compute canonical loadings
cc2 <- comput(psych, acad, cc1)
# display canonical loadings
cc2[3:6]


CCA::matcor(psych, acad)
```


interpret results: cc1的结果和回归系数的情况是很像的，i.e., for the variable read, a one unit increase in reading leads to a .0446 decrease in the first canonical variate of set 2 when all of the other variables are held constant.
The above correlations are between observed variables and canonical variables which are known as the canonical loadings. 



Canonical dimensions, also known as canonical variates, are latent variables that are analogous to factors obtained in factor analysis. For this particular model there are three canonical dimensions of which only the first two are statistically significant. For statistical test we use R package "CCP".
```{r}
# tests of canonical dimensions
rho <- cc1$cor
## Define number of observations, number of variables in first set, and number of variables in the second set.
n <- dim(psych)[1]
p <- length(psych)
q <- length(acad)

## Calculate p-values using the F-approximations of different test statistics:
p.asym(rho, n, p, q, tstat = "Wilks")

p.asym(rho, n, p, q, tstat = "Hotelling")
```



When the variables in the model have very different standard deviations, the standardized coefficients allow for easier comparisons among the variables. Next, we’ll compute the standardized canonical coefficients.
```{r}
# standardized psych canonical coefficients diagonal matrix of psych sd's
s1 <- diag(sqrt(diag(cov(psych))))
s1 %*% cc1$xcoef

# standardized acad canonical coefficients diagonal matrix of acad sd's
s2 <- diag(sqrt(diag(cov(acad))))
s2 %*% cc1$ycoef
```







