---
title: "Hidden Markov Models"
author: "liuc"
date: "11/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hidden Markov Models

```{r}
library(depmixS4)
library(easystats)
```

A Hidden Markov model is a tool for representing probability distribution over a sequence of observations. It assumes that the observation at time t was generated by some process whose state is hidden from the observer.
通过观察observation的分布，推测hidden state。Transition probabilities and Emission Probabilities分别表示 observation和hidden的事件发生概率。
先验概率的数据来自哪呢？


隐马尔可夫在生物信息中的应用主要是用于序列性数据的分析。比如DNA测序后call mutation。
```{r}
data(speed)

# HMM with depmix
mod <- depmix(response = rt ~ 1, 
              data = speed, 
              nstates = 2, 
              family = gaussian()) # use gaussian() for normally distributed data

fit.mod <- fit(mod)
fit.mod
summary(fit.mod, which = "transition")

# predict the states by estimating the posterior
est.states <- posterior(fit.mod)
head(est.states)


```


```{r}

```


