---
title: "PCA_UMAP_tidymodels"
author: "liuc"
date: '2022-05-27'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PCA_UMAP_tidymodels

https://juliasilge.com/blog/cocktail-recipes-umap/

在`recipe`包中提供了多种特征工程的方便的函数，其也是包装了一些其他函数的功能，在使用时虽则简单了很多，但还是需要理解其所使用的后端的函数，以及对应参数的调整。
`tidymodels`还有提供多个`recipe`的扩展包，如以下的`embed`包。



```{r}
library(tidymodels)
library(usemodels)
library(finetune)
library(embed) # recipe external package
```



```{r}
boston_cocktails <- readr::read_csv("./datasets/boston_cocktails.csv")

boston_cocktails %>%
  count(ingredient, sort = TRUE)


cocktails_parsed <- boston_cocktails %>%
  mutate(
    ingredient = str_to_lower(ingredient),
    ingredient = str_replace_all(ingredient, "-", " "),
    ingredient = str_remove(ingredient, " liqueur$"),
    ingredient = str_remove(ingredient, " (if desired)$"),
    ingredient = case_when(
      str_detect(ingredient, "bitters") ~ "bitters",
      str_detect(ingredient, "lemon") ~ "lemon juice",
      str_detect(ingredient, "lime") ~ "lime juice",
      str_detect(ingredient, "grapefruit") ~ "grapefruit juice",
      str_detect(ingredient, "orange") ~ "orange juice",
      TRUE ~ ingredient
    ),
    measure = case_when(
      str_detect(ingredient, "bitters") ~ str_replace(measure, "oz$", "dash"),
      TRUE ~ measure
    ),
    measure = str_replace(measure, " ?1/2", ".5"),
    measure = str_replace(measure, " ?3/4", ".75"),
    measure = str_replace(measure, " ?1/4", ".25"),
    measure_number = parse_number(measure),
    measure_number = if_else(str_detect(measure, "dash$"),
      measure_number / 50,
      measure_number
    )
  ) %>%
  add_count(ingredient) %>%
  filter(n > 15) %>%
  select(-n) %>%
  distinct(row_id, ingredient, .keep_all = TRUE) %>%
  na.omit()

cocktails_parsed

cocktails_df <- cocktails_parsed %>%
  select(-ingredient_number, -row_id, -measure) %>%
  pivot_wider(names_from = ingredient, values_from = measure_number, values_fill = 0) %>%
  janitor::clean_names() %>%
  na.omit()

cocktails_df # 行为特征、列为sample的矩阵
```


#### PCA

The step_pca() function uses stats::prcomp() under the hood.

在`tidymodels`中，`recipe`包即可提供PCA等非监督的降维方法。不过所提供的参数和分析还是不如独立的分析。
PCA作为一种降维方法，
```{r}
pca_rec <- recipe(~., data = cocktails_df) %>%
  update_role(name, category, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

# Before using prep() these steps have been defined but not actually run or implemented. The prep() function is where everything gets evaluated.
pca_prep <- prep(pca_rec)

pca_prep
```



```{r}
tidied_pca <- tidy(pca_prep, 2) # 2为recipe中的第二步的step
tidied_pca

tidied_pca %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component, nrow = 1) +
  labs(y = NULL)
```


This is the component that explains the most variation in drinks
```{r}
library(tidytext)

tidied_pca %>%
  filter(component %in% paste0("PC", 1:4)) %>%
  group_by(component) %>%
  top_n(8, abs(value)) %>%
  ungroup() %>%
  mutate(terms = reorder_within(terms, abs(value), component)) %>%
  ggplot(aes(abs(value), terms, fill = value > 0)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y") +
  scale_y_reordered() +
  labs(
    x = "Absolute value of contribution",
    y = NULL, fill = "Positive?"
  )


juice(pca_prep) %>%
  ggplot(aes(PC1, PC2, label = name)) +
  geom_point(aes(color = category), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)
```

绘制PCA的散点图。
```{r}
bake(pca_prep, new_data = NULL) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = category), alpha = 0.7, size = 2) +
  # geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL) +
  theme_bw()

```



#### UMAP

We included UMAP since it can be projected on new data. tSNE can't (at least, the last time that I looked).

```{r}
umap_rec <- recipe(~., data = cocktails_df) %>%
  update_role(name, category, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_umap(all_predictors())

umap_prep <- prep(umap_rec)

umap_prep
```



how the cocktails are distributed in the plane of the first two UMAP components.
```{r}
juice(umap_prep) %>%
  ggplot(aes(umap_1, umap_2, label = name)) +
  geom_point(aes(color = category), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward", family = "IBMPlexSans") +
  labs(color = NULL)
```




