---
title: "pharmacokinetics modelling"
author: "liuc"
date: "1/12/2022"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## pharmacokinetics modelling by R

> http://sia.webpopix.org/pharmacokinetics.html
> http://sia.webpopix.org/nlme.html
> 熟悉nlmixr2包的文档
> https://www.pkpd-info.com/PK_introduction.php


Once a drug is administered, we usually describe subsequent processes within the organism by the pharmacokinetics (PK) process known as ADME: absorption, distribution, metabolism, excretion.
具体的关于PK的一些描述可以参见第四个链接。

将人体简化为depot compartment。PK compartment models assume that drug concentration is perfectly homogeneous in each compartment of the body at all times. 

PK models do not describe how a drug is administered. Administered doses are source terms, i.e., inputs that dynamically modify the state of a system. This is important to note because it means that the same PK model can be used for different dose regimens.

ordinary differential equations (ODEs):

PK模型有很多，但非线性混合模型是PK建模中常用的模型。下面以一个测试数据集进行初步的学习，此处做一笔记。

k : rate constants (kinetic) --> movement from 'x' to 'y'
Ka: absorption constant
Vc: distribution compartment (central)
Q: rate constant between Vc and Vp
Vp: distribution compartment (periph.)
CL: elimination constant (clearance)
Described by differential equations of change in compartment (A) over time (T): dA/dT

```{r, include=FALSE}
# library(saemix)
# library(nlmixr) # 主要学习的R包，不过和2的版本不兼容，二者先library一个
library(nlmixr2)
library(rxode2)
library(ggPMX)
# library(ubiquity)
library(PKNCA)
library(xpose)

# library('nonmem2R')

```


### single-dose

 fitting a simple one-compartment PK model use `nlmixr2` package.
 
 one-compartment指的是，
 
 下面的模型中，和在使用`saemix`包时是一致的，包括模型和起始值。`nlmixr2`包的语法和ini的命名系统是模仿的NONMEM软件。在ini中设置值时一般的是一个值，或者三个值（c(lower, est, upper)）,`~`是方差或相关性，`<-`是参数赋值,`#`后面的内容是 label的内容会在Parameter展现，也可以用label函数。
 
 
其中，model部分的意思是：A mathematical model which best describes the data

theo_sd是常见的theophylline数据集。在`nlmixr2`中数据

```{r}
#  ordinary differential equations (ODEs) model
one.compartment.saem <- function() {
    ini({
        tka <- .5   # Log Ka
        tcl <- -3.2 # Log Cl
        tv <- -1    # Log V
        eta.ka ~ 1 ## BSV Ka
        eta.cl ~ 2
        eta.v ~ 1
        add.err <- 0.1 # residual variability
    })
    model({
        ka <- exp(tka + eta.ka)
        cl <- exp(tcl + eta.cl)
        v <- exp(tv + eta.v)
        d/dt(depot) = -ka * depot
        d/dt(center) = ka * depot - cl / v * center
        cp = center / v 
        cp ~ add(add.err)
    })
}

# 数据集相关的变量啥也不用指定吗？？
fit <- nlmixr2(one.compartment.saem, 
               data = theo_sd, 
               est="saem" # nlmixr2est::nlmixr2AllEst()
               )

```

```{r}
print(fit)
```


base

```{r}
plot(fit)
```

除上述通过`base`包得到的诊断图之外，还提供了`xpose`包的使用，`xpose`包是aims to reduce the post processing burden and improve diagnostics commonly associated the development of non-linear mixed effect models.

```{r}
library(xpose.nlmixr2)

xpdb <- xpose_data_nlmixr(fit) # 此object可以使用xpose包的函数

xpose::dv_vs_ipred(xpdb)
```







本文主要从处方前研究、剂型开发、食物效应、预测人体药代动力学以及仿制药的生物等效性研究等方面介绍了PBPK 建模在药剂学领域中的应用

```{r}

```




### use PKNCA 包

















